---
title: "Wicked Problems Proposal"
---
```{r include=FALSE}
library(tidyverse)
library(haven)
library(readr)
library(readxl)
demog <- read_csv("data/demog.csv", show_col_types = F)
```

# Introduction

Wicked Problems are defined as complex challenges that lack straightforward solutions, often causing disagreement amongst multiple stakeholders pertaining to the best policy solution for these proposed problems. While the term is normally seen as bad due to their complexity as well the negative outcomes given some of these issues, they are not  inherently bad, in fact depending on who is asked, these problems might not exist. In a pilot study done during the semester I asked 25 individuals to rate the connectivity of 35 concepts that were generated by my fellow classmates. In doing this I hope to understand two things, the perception of these concepts as problems in the United States, and how individuals perceive the connection between these concepts. 


## Research questions:
What is the perception of these concepts as problems in the United states based on the respondents current definition of the concepts?

What are the perceived connections between these concepts and how can they be grouped together and summed up into an overarching theme.

## Hypothesis
  I am expecting to find imperalism as a key influence in one of the groups, I also expect economy and the gov't structure to be in the same group.
  
  The largest problem im expecting is Racism and the Economy as these are major issues in the news in the past couple of years between the BLM protest, January 6th and inflation after the pandemic.

## Demgraphics
### politcal party based on age and region
```{r}
counts <- table(demog$age, demog$`party _d`, demog$region)

# Reshape data for ggplot
counts <- as.data.frame(counts)
names(counts) <- c('age', 'party_id', 'region','count')

# Plotting
ggplot(counts, aes(x = factor(age), y = count, fill = party_id)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(x = 'Age', y = 'Counts', fill = 'Party ID') +
  scale_fill_manual(values = c('pink', 'skyblue', 'lightgreen', 'orange')) +
  ggtitle('Party IDs Across Age Groups') +
  theme_minimal()



ggplot(counts, aes(x = factor(region), y = count, fill = party_id)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(x = 'Age', y = 'Counts', fill = 'Party ID') +
  scale_fill_manual(values = c('pink', 'skyblue', 'lightgreen', 'orange')) +
  ggtitle('Party IDs Across Age Groups') +
  theme_minimal()

```
### Problem rating by party idea
```{r}
pivot_demog <- pivot_longer(demog, cols = Imperialism:Eurocentrism, names_to = "Concept", values_to = "Ratings")

# Plotting
ggplot(pivot_demog, aes(x = Concept, y = Ratings, fill = `party _d`)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(x = 'Concepts', y = 'Ratings', fill = 'Respondents') +
  ggtitle('Ratings of Concepts by Respondents') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
### Overall/Average Rating
```{r}

average_ratings <- pivot_demog %>%
  group_by(Concept) %>%
  summarise(Average_Rating = mean(Ratings))

# Plotting
library(ggplot2)

ggplot(average_ratings, aes(x = Concept, y = Average_Rating)) +
  geom_bar(stat = 'identity', fill = 'skyblue', width = 0.5) +
  geom_text(aes(label = round(Average_Rating, 2)), vjust = -0.5, size = 4, color = "black") +
  labs(x = 'Concepts', y = 'Average Ratings') +
  ggtitle('Average Ratings of Concepts') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```

## Methods:
Survey Analysis -   
