---
title: "Wicked Problems Proposal"
---

```{r include=FALSE}
library(tidyverse)
library(haven)
library(readr)
library(readxl)
demog <- read_csv("data/demog.csv", show_col_types = F)
```

# Introduction

Wicked Problems are defined as complex challenges that lack straightforward solutions, often causing disagreement among multiple stakeholders pertaining to the best policy solution for these proposed problems. While the term is normally seen as bad due to their complexity as well the negative outcomes given some of these issues, they are not inherently bad, in fact depending on who is asked, these problems might not exist. In a pilot study done during the semester I asked 25 individuals to rate the connectivity of 35 concepts that were generated by my fellow classmates. In doing this I hope to understand two things, the perception of these concepts as problems in the United States, and how individuals perceive the connection between these concepts.

## Research questions:

What is the perception of these concepts as problems in the United states based on the respondents current definition of the concepts?

What are the perceived connections between these concepts and how can they be grouped together and summed up into an overarching theme.

## Hypothesis

I am expecting to find imperialism as a key influence in one of the groups, I also expect economy and the gov't structure to be in the same group.

The largest problem im expecting is Racism and the Economy as these are major issues in the news in the past couple of years between the BLM protest, January 6th and inflation after the pandemic.

## Demgraphics

#### political party based on age

```{r}
counts <- table(demog$age, demog$`party _d`, demog$region)

# Reshape data for ggplot
counts <- as.data.frame(counts)
names(counts) <- c('age', 'party_id', 'region','count')

# Plotting
ggplot(counts, aes(x = factor(age), y = count, fill = party_id)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(x = 'Age', y = 'Counts', fill = 'Party ID') +
  scale_fill_manual(values = c('pink', 'skyblue', 'lightgreen', 'orange')) +
  ggtitle('Party IDs Across Age Groups') +
  theme_minimal()
```

#### political party based on region

```{r}


ggplot(counts, aes(x = factor(region), y = count, fill = party_id)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(x = 'Age', y = 'Counts', fill = 'Party ID') +
  scale_fill_manual(values = c('pink', 'skyblue', 'lightgreen', 'orange')) +
  ggtitle('Party IDs Across Age Groups') +
  theme_minimal()

```

## Concept ratings

Survey Analysis - Based on survey results the respondents rated how much of a problem they believed each concept to be on a scale of 1 to 5. From there I averaged the results for each concept and plotted them on a bar graph. \#### concepts rating by party idea

```{r}
pivot_demog <- pivot_longer(demog, cols = Imperialism:Eurocentrism, names_to = "Concept", values_to = "Ratings")

# Plotting
ggplot(pivot_demog, aes(x = Concept, y = Ratings, fill = `party _d`)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(x = 'Concepts', y = 'Ratings', fill = 'Respondents') +
  ggtitle('Ratings of Concepts by Respondents') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Overall/Average Problem Rating

```{r}

average_ratings <- pivot_demog %>%
  group_by(Concept) %>%
  summarise(Average_Rating = mean(Ratings))

# Plotting
library(ggplot2)

ggplot(average_ratings, aes(x = Concept, y = Average_Rating)) +
  geom_bar(stat = 'identity', fill = 'skyblue', width = 0.5) +
  geom_text(aes(label = round(Average_Rating, 2)), vjust = -0.5, size = 3, color = "black") +
  labs(x = 'Concepts', y = 'Average Ratings') +
  ggtitle('Average Ratings of Concepts') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability
```

##Concept Grouping

I used the average clustering method to group the concepts together based on their degree strength. Its worth noting that I used tie count as both degree strength and weight. I used Tnet package to allow me to choose which measurement should be prioritized and for this project I prioritized degree strength to identify centrality, specifically out-degree strength.

from there I partitioned the clustering at height 5, which gave me 9 categorizes which I developed my block model on and then created sub graphs based on structural equivalence. Once I had the subgraphs Identified the most central node of the 9 groups as the node with the greatest out degree, as I got 9 directed graphs.

### Categories:

#### Power Dynamics:

Topics related to the exertion, distribution, and dynamics of power in society, politics, and economics. This category encompasses topics such as imperialism, racism, colonization, capitalism, and other forms of power dynamics.

#### Social Infrastructure:

Topics related to the foundational elements of society that support its functioning, including housing, education, infrastructure, and criminal justice.

#### Economic & Government Structure:

Topics concerning the economic system, government structures, and their interactions, including economy, income/wealth, corporate wealth, federal/government structure, and related aspects.

#### Regulation & Control:

Topics pertaining to regulatory measures, control mechanisms, and governance practices, such as gun control, banking regulations, voting access, and other forms of regulation.

#### Global Challenges:

Topics addressing global issues and challenges that transcend national boundaries, including climate change, child poverty, nationalizing, digital literacy, and other global challenges. 

#### Social & Global Issues: 

Topics encompassing social and global issues that impact societies worldwide, such as AI, immigration, misinformation, pollution & plastic, and indigenous sovereignty.

#### Trust & Rights:

Topics revolving around trust, rights, and societal values, including weakening of the global West, public trust, homeless rights, welfare systems, and civil rights.

#### Media & Governance:

Topics concerning media, journalism, governance, and their roles in society, including journalism/media, foreign policy, insider trading, generational wealth, and Eurocentrism. 

#### Cultural Awareness & Equity: 
Topics focusing on cultural awareness, equity, and diversity, such as cultural awareness & equity and Eurocentrism.
